---
title: "Initial Exploration of the Arrival Time Survey"
author: "Akansha Vashisth, Ian Flores Siaca, Rachel K. Riggs, Milos Milic"
date: '2019-04-06'
output: github_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(warning=FALSE, message = FALSE)
library(tidyverse)
library(lubridate)
library(skimr)
library(pander)
```

### Survey question

**How does distance from campus influence arrival time to lectures?**

### Collected data

```{r}
# Initial survey data
initial_survey <- read_csv('data/Arrival_Time_April_4_2019_08_20.csv', col_type = cols())

head(initial_survey)
```


### Cleaning data

```{r}
clean_survey_sep_days <- initial_survey %>%
  select(contains('Q')) %>%
  slice(-c(1:2)) %>%
  rename('distance_km' = 'Q1',
         'mw_arrival' = 'Q3',
         'tt_arrival' = 'Q4',
         'mode_of_transport' = 'Q5') %>% 
  mutate(mw_arrival = hm(mw_arrival),
         mw_arrival = hour(mw_arrival)*60 + minute(mw_arrival),
         mw_arrival = if_else(mw_arrival < 400, 540 + mw_arrival, mw_arrival),
         mw_arrival = mw_arrival - 9*60) %>%
  mutate(tt_arrival = hm(tt_arrival),
         tt_arrival = hour(tt_arrival)*60 + minute(tt_arrival),
         tt_arrival = if_else(tt_arrival < 430, 570 + tt_arrival, tt_arrival),
         tt_arrival = tt_arrival - 9.5*60) %>%
  mutate(mode_of_transport = fct_recode(mode_of_transport, 
                                        driving = '1',
                                        transit = '2',
                                        walking = '3',
                                        cycling = '4')) %>%
    mutate(distance_km = as.numeric(distance_km))


clean_survey_mw <- clean_survey_sep_days %>% 
  select(-c(tt_arrival)) %>% 
  rename(arrival = mw_arrival)

clean_survey_tt <- clean_survey_sep_days %>% 
  select(-c(mw_arrival)) %>% 
  rename(arrival = tt_arrival)

clean_survey_all_days <- bind_rows(clean_survey_mw, clean_survey_tt)
```


### Cleaned data

```{r}
head(clean_survey_sep_days)
head(clean_survey_all_days)
```


### Variables in the data
1. distance_km: Distance travelled in kilometers. This is a numeric variable.
2. mw_arrival: Arrival time in class on Mondays and Wednesdays. This is a numeric variable.
3. tt_arrival: Arrival time in class on Tuesdays and Thursdays. This is a numeric variable.
4. mode_of_transport: Mode of transportation used to travel. This is a categorical variable.


### Table of Summary Statistics

#### General Summary

```{r, include = FALSE}
panderOptions('knitr.auto.asis', FALSE)
skim_with(numeric = list(hist = NULL), factor = list(ordered = NULL))
```

```{r}
skim(clean_survey_sep_days) %>% 
    kable()
```

#### Grouped by Mode of Transport

```{r, include = FALSE}
skim_with(numeric = list(hist = NULL, missing = NULL, complete = NULL))
```

```{r}
clean_survey_sep_days %>%
    group_by(mode_of_transport) %>%
    skim() %>%
    kable()
```

### Exploratory Data Analysis


```{r}
### put plots for MW and TT here


clean_survey_sep_days %>% 
  ggplot(aes(x = mw_arrival)) +
  geom_density() + 
  facet_wrap(~ mode_of_transport)+
  labs(x="Arrival time", y="Frequency", title = "Count of student arrival time on Monday Wednesday")


clean_survey_sep_days %>% 
  ggplot(aes(x = tt_arrival)) +
  geom_density() + 
  facet_wrap(~ mode_of_transport)+
  labs(x="Arrival time", y="Frequency", title = "Count of student arrival time on Tuesday Thursday")
```



```{r}
### violin plot for all days, faceted by mode of transport
clean_survey_mw %>% 
  ggplot(aes(x = mode_of_transport, y = arrival)) +
  geom_violin() +
  geom_jitter(width = .2, alpha = .3)+
  labs(y="Arrival time", x="Mode of transport", title = "Mode of transport v/s arrival time on (Monday Wednesday)")

clean_survey_tt %>% 
  ggplot(aes(x = mode_of_transport, y = arrival)) +
  geom_violin() +
  geom_jitter(width = .2, alpha = .3)+
  labs(y="Arrival time", x="Mode of transport", title = "Mode of transport v/s arrival time on (Tuesday Thursday)")



```




```{r}
### scatterplot for all days grouped together, faceted by mode of transport
clean_survey_sep_days %>% 
  ggplot(aes(x = mw_arrival, y = distance_km)) +
  geom_point() +
  facet_wrap(~ mode_of_transport) +
  labs(x="Arrival time", y="Distance(in kms)", title = "Arrival time v/s distance")
```


```{r}
### density plot for all days, faceted by mode of transport
clean_survey_all_days %>% 
  ggplot(aes(x = arrival)) +
  geom_density() + 
  facet_wrap(~ mode_of_transport)+
  labs(x="Arrival time", y="Frequency", title = "Count of student arrival time")
```


```{r}
### boxplot for all days, faceted by mode of transport
clean_survey_all_days %>% 
  ggplot(aes(x = mode_of_transport, y = arrival)) +
  geom_boxplot()+
  labs(y="Arrival time", x="Mode of transport", title = "Mode of transport v/s arrival time")
```

```{r}
### violin plot for all days, faceted by mode of transport
clean_survey_all_days %>% 
  ggplot(aes(x = mode_of_transport, y = arrival)) +
  geom_violin() +
  geom_jitter(width = .2, alpha = .3)+
  labs(y="Arrival time", x="Mode of transport", title = "Mode of transport v/s arrival time")
```

### Conclusion
(open to all)
> Based on the EDA
