---
title: "Survey Analysis Arrival Time"
author: "Akansha Vashisth, Ian Flores Siaca, Rachel K. Riggs, Milos Milic"
date: '2019-04-13'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Libraries

```{r, warning=FALSE}
library(tidyverse)
library(tidybayes)
library(brms)
library(broom)
library(knitr)
library(gridExtra)
```

## Load the data

```{r}
clean_survey_all_days <- read_csv('https://raw.githubusercontent.com/UBC-MDS/survey_arrival_time/master/data/clean_survey_responses_all_days.csv')
clean_survey_sep_days <- read_csv('https://raw.githubusercontent.com/UBC-MDS/survey_arrival_time/master/data/clean_survey_responses_sep_days.csv')
```

# Survey question

**How does distance from campus influence arrival time to lectures?**

We conducted an observational study to explore if there is a relationship between the distance lived from class and arrival time. We also wanted to test a potential confounder for this relationship which is the mode of transportation a student takes to class.

# Methods

## Survey study design

We asked our survey respondents to answer the following questions:

- How far from Hugh Dempster do you live in kilometers via the mode of transport you use(a google maps link was provided to help with the distance estimation)?

- What time do you typically arrive at Hugh Dempster on Mondays and Wednesdays?
please enter in the format hh:mm

- What time do you typically arrive at Hugh Dempster on Mondays and Wednesdays?
please enter in the format hh:mm

- What is your typical mode of transit? (drive, public transit, walk, or bike)

## Data collection methods

Data was collected by this [survey](https://ubc.ca1.qualtrics.com/jfe/form/SV_eo1whP0fPfWPCw5) hosted by Qualtrics. The survey had 56 participants from the MDS students 2018-2019 cohort and responses were anonymized.

## Analysis methods

We performed initial [exploratory data analysis](https://github.com/UBC-MDS/survey_arrival_time/blob/v2.0/milestone2.md) on our data. 

To analyze the data, we consider 3 groups:

- all days grouped together

- Mondays and Wednesdays together

- Tuesdays and Thursdays together

Below we fit a linear regression using the distance as the predictor variable and the arrival time as the response variable. We compare this linear regression model with a null model through an ANOVA test. To validate the estimates of the frequentist approach, given the possibility of a small sample size, we use a Bayesian linear regression. After this, we will move on to using the mode of transportation as a confounder variable and fit a linear regression model with these variables. We will compare this model with the null model through an ANOVA test, and again validate using a Bayesian linear regression.

# EDA

```{r}
plot1 <- clean_survey_all_days %>%
  ggplot(height = 17 , width = 2) +
  geom_histogram(aes(x=distance_km)) + 
  theme(axis.title=element_text(size=10),
        plot.title = element_text(size = 10, face = "bold")) +  
  labs(y= "Frequency", x = "Distance (in km)", title = "Distribution of Distance Lived From Campus (km)") 

plot2 <- clean_survey_all_days %>%
  ggplot() +
  geom_histogram(aes(x=arrival)) +  
  theme(axis.title=element_text(size=10),
        plot.title = element_text(size = 10, face = "bold")) +  
  labs(y= "Frequency", x = "Arrival time", title = "Distribution of Arrival Time (minutes)") 

grid.arrange(plot1, plot2, ncol=2)
```

We can see that the majority of students live within 15 kilometers of campus, and that the majority arrive within 30 mintues before and 30 minutes after the start of the lecture.

```{r}
plot3 <- clean_survey_all_days %>%
  ggplot() +
  geom_bar(aes(x=mode_of_transport)) +
  theme(axis.title=element_text(size=10),
        plot.title = element_text(size = 10, face = "bold")) +  
  labs(y= "Frequency", x = "Mode of transport", title = "Number of MDS Students using different modes of transport") 
plot3
```

Comparing modes of transit, we see that transit is the most common form of transportation, while cycling is the least.

Does there appear to be a relationship between arrival time and distance lived from campus? An intial plot shows us:

```{r}
plot4 <- clean_survey_all_days %>%
  ggplot(aes(x = arrival, y = distance_km)) +
  geom_point() +
  theme(axis.title=element_text(size=10),
        plot.title = element_text(size = 10, face = "bold")) +  
  labs(x = "Arrival time (minutes)", y = "Distance (km)", title = "Arrival Time vs Distance Lived From Campus") 
plot4
```

It's difficult to discern from this plot so we will further explore this question in the analysis below.

```{r}
plot5 <- clean_survey_sep_days %>% 
  ggplot(aes(x = mw_arrival)) +
  geom_density(aes(fill = "salmon", color = "salmon"), alpha = .3) + 
  geom_density(aes(x = tt_arrival, fill = "#00BFC4", color = "#00BFC4"), alpha = .3) +
  facet_wrap(~ mode_of_transport) +
  theme(axis.title=element_text(size=10),
        plot.title = element_text(size = 10, face = "bold")) +
  labs(x="Arrival time", y="Frequency", title = "Density of student arrival time on Mon & Wed vs Tues & Thurs") +
  guides(color = FALSE) +
  scale_fill_identity(name = "days",
                       breaks = c("salmon", "#00BFC4"),
                       labels = c("Mon-Wed", "Tues-Thurs"),
                       guide = "legend")
plot5
```

Comparing this distribution of arrival times on Mondays and Wednesday vs Tuesdays and Thursdays reveals that there is only a slight difference in arrival time when the lectures start at 09:00 compared when the lecture starts at 09:30. It is most noticeable for students taking transit.

# Analysis and results

## Without Confounders

### Distance and Arrival Time (Overall)

```{r}
# Frequentist
fit_all <- lm(arrival ~ distance_km, data = clean_survey_all_days)

tidy(fit_all) %>% 
    kable()
```



```{r}
# Shows the Frequentist 0.95 confidence intervals

confint_tidy(fit_all, conf.level = 0.95)

```


As per frequentist approach there is an association between distance and arrival time with a confidence interval of 95% (-1.604, -0.257).

For every increase of 1 km lived from campus, the expected change in arrival time is early by almost 1 minute.


```{r}
#Frequentist plot

ggplot(clean_survey_all_days, aes(x = arrival, y = distance_km)) + 
  geom_point() +
  xlab("arrival time relative to start of class") +
  ylab("distance in km") +
  ggtitle("Distance vs arrival time plot") +
  stat_smooth(method = "lm", col = "red")

```





```{r, include=FALSE, cache=TRUE}
#Bayes Model

fit_all_bayes <- brm(arrival ~ distance_km, data = clean_survey_all_days, iter = 5000, cores = -1)
```

```{r, cache=TRUE}
# Returns the 95% estimates
fit_all_bayes %>%
    gather_draws(b_Intercept, b_distance_km, sigma) %>%
    median_qi() %>%
    kable()
```

As per bayesian approach (Ian could you please take over from here)

```{r, cache=TRUE}
#Plots the Baysean model
fit_all_bayes %>%
    gather_draws(b_Intercept, b_distance_km) %>%
    ggplot(aes(.value)) +
    geom_density() +
    facet_wrap(~ .variable, scales = 'free') +
    labs(x = 'Value',
         title = 'Bayesian All Days Model')
```

### Distance and Arrival Time (Monday & Wednesday)

```{r}
# Frequentist Model
fit_mw <- lm(mw_arrival ~ distance_km, data = clean_survey_sep_days)

tidy(fit_mw) %>% 
    kable()
```



```{r}
# Shows the Frequentist 0.95 confidence intervals

confint_tidy(fit_mw, conf.level = 0.95)

```


As per frequentist approach there is an association between distance and arrival time with a confidence interval of 95% (-1.76, 0.017) for Monday and Wednesday classes(9 am start).



```{r}
#Frequentist plot

ggplot(clean_survey_sep_days, aes(x = mw_arrival, y = distance_km)) + 
  geom_point() +
  xlab("arrival time relative to start of class (Monday and Wednesday)") +
  ylab("distance in km") +
  ggtitle("Distance vs arrival time plot(Monday and Wednesday)") +
  stat_smooth(method = "lm", col = "red")

```





```{r, include=FALSE, cache=TRUE}
#Bayes Model
fit_mw_bayes <- brm(mw_arrival ~ distance_km, data = clean_survey_sep_days, iter = 5000, cores = -1)
```

```{r, cache=TRUE}
# Returns the 95% estimates
fit_mw_bayes %>%
    gather_draws(b_Intercept, b_distance_km, sigma) %>%
    median_qi() %>%
    kable()
```

```{r, cache=TRUE}
# Plots the Bayes Model
fit_mw_bayes %>%
    gather_draws(b_Intercept, b_distance_km) %>%
    ggplot(aes(.value)) +
    geom_density() +
    facet_wrap(~ .variable, scales = 'free') +
    labs(x = 'Value',
         title = 'Bayesian Monday & Wednesday Model')
```

### Distance and Arrival Time (Tuesday and Thursdays)

```{r}

#Frequentist model
fit_tt <- lm(tt_arrival ~ distance_km, data = clean_survey_sep_days)

tidy(fit_tt) %>% 
    kable()
```




```{r}
# Shows the Frequentist 0.95 confidence intervals

confint_tidy(fit_tt, conf.level = 0.95)

```


As per frequentist approach there is an association between distance and arrival time with a confidence interval of 95% (-2.1, 0.0677) for Tuesday and Thursday classes(9:30 am start).



```{r}
#Frequentist plot

ggplot(clean_survey_sep_days, aes(x = tt_arrival, y = distance_km)) + 
  geom_point() +
  xlab("arrival time relative to start of class (Tuesday and Thursday)") +
  ylab("distance in km") +
  ggtitle("Distance vs arrival time plot(Tuesday and Thursday)") +
  stat_smooth(method = "lm", col = "red")

```






```{r, include=FALSE, cache=TRUE}
#Bayes Model
fit_tt_bayes <- brm(tt_arrival ~ distance_km, data = clean_survey_sep_days, iter = 5000, cores = -1)
```

```{r, cache=TRUE}
# Returns the 95% estimates

fit_tt_bayes %>%
    gather_draws(b_Intercept, b_distance_km, sigma) %>%
    median_qi() %>%
    kable()
```

```{r, cache=TRUE}
# Plots the Bayes Model
fit_tt_bayes %>%
    gather_draws(b_Intercept, b_distance_km) %>%
    ggplot(aes(.value)) +
    geom_density() +
    facet_wrap(~ .variable, scales = 'free') +
    labs(x = 'Value',
         title = 'Bayesian Tuesday & Thursday Model')
```

## With Confounders

### Distance and Arrival Time (Overall)

```{r}
#Frequentist approach
fit_all_transp <- lm(arrival ~ distance_km + mode_of_transport, data = clean_survey_all_days)

tidy(fit_all_transp) %>% 
    kable()
```




```{r}
# Shows the Frequentist 0.95 confidence intervals

confint_tidy(fit_all_transp, conf.level = 0.95)

```



```{r}
#Frequentist plot

ggplot(clean_survey_all_days, aes(x = arrival, y = distance_km)) + 
  geom_point() +
  facet_wrap(~mode_of_transport, nrow = 4) +
  xlab("arrival time relative to start of class ") +
  ylab("distance in km") +
  ggtitle("Distance vs arrival time plot") +
  stat_smooth(method = "lm", col = "red")

```









```{r, include=FALSE, cache=TRUE}
#Bayes model

fit_all_bayes_transp <- brm(arrival ~ distance_km + mode_of_transport, data = clean_survey_all_days, iter = 5000, cores = -1)
```

```{r, cache=TRUE}
# Returns the 95% estimates
fit_all_bayes_transp %>%
    gather_draws(b_Intercept, b_distance_km, sigma) %>%
    median_qi() %>%
    kable()
```

```{r, cache=TRUE}
#Bayes model plot

fit_all_bayes_transp %>%
    gather_draws(b_Intercept, b_distance_km) %>%
    ggplot(aes(.value)) +
    geom_density() +
    facet_wrap(~ .variable, scales = 'free') +
    labs(x = 'Value',
         title = 'Bayesian All Days Model w Confounder')
```

### Distance and Arrival Time (Monday & Wednesday)

```{r}

#Frequentist
fit_mw_transp <- lm(mw_arrival ~ distance_km + mode_of_transport, data = clean_survey_sep_days)

tidy(fit_mw_transp) %>% 
    kable()
```




```{r}
# Shows the Frequentist 0.95 confidence intervals

confint_tidy(fit_mw_transp, conf.level = 0.95)

```



```{r}
#Frequentist plot

ggplot(clean_survey_sep_days, aes(x = mw_arrival, y = distance_km)) + 
  geom_point() +
  facet_wrap(~mode_of_transport, nrow = 4) +
  xlab("arrival time relative to start of class (Monday and Wednesday) ") +
  ylab("distance in km") +
  ggtitle("Distance vs arrival time plot(Monday and Wednesday)") +
  stat_smooth(method = "lm", col = "red")

```






```{r, include=FALSE, cache=TRUE}
#Bayes model

fit_mw_bayes_transp <- brm(mw_arrival ~ distance_km + mode_of_transport, data = clean_survey_sep_days, iter = 5000, cores = -1)
```

```{r, cache=TRUE}
#Returns 95% estimates

fit_mw_bayes_transp %>%
    gather_draws(b_Intercept, b_distance_km, sigma) %>%
    median_qi() %>%
    kable()
```

```{r, cache=TRUE}
#Plots the Bayes model

fit_mw_bayes_transp %>%
    gather_draws(b_Intercept, b_distance_km) %>%
    ggplot(aes(.value)) +
    geom_density() +
    facet_wrap(~ .variable, scales = 'free') +
    labs(x = 'Value',
         title = 'Bayesian Monday & Wednesday Model w Confounder')
```

### Distance and Arrival Time (Tuesday and Thursdays)

```{r}
#Frequentist
fit_tt_transp <- lm(tt_arrival ~ distance_km + mode_of_transport, data = clean_survey_sep_days)

tidy(fit_tt_transp) %>% 
    kable()
```






```{r}
# Shows the Frequentist 0.95 confidence intervals

confint_tidy(fit_tt_transp, conf.level = 0.95)

```

The confidence intervals for all of the modes of transport are quite different cycling(-2.36 and 0.523), driving(-57.4 and 26.3), transit (-30.2 and 37.6), and walking (-38.6 and 32.8)



```{r}
#Frequentist plot

ggplot(clean_survey_sep_days, aes(x = tt_arrival, y = distance_km)) + 
  geom_point() +
  facet_wrap(~mode_of_transport, nrow = 4) +
  xlab("arrival time relative to start of class (Tuesday and Thursday) ") +
  ylab("distance in km") +
  ggtitle("Distance vs arrival time plot (Tuesday and Thursday)") +
  stat_smooth(method = "lm", col = "red")

```







```{r, include=FALSE, cache=TRUE}
fit_tt_bayes_transp <- brm(tt_arrival ~ distance_km + mode_of_transport, data = clean_survey_sep_days, iter = 5000, cores = -1)
```

```{r, cache=TRUE}
fit_tt_bayes_transp %>%
    gather_draws(b_Intercept, b_distance_km, sigma) %>%
    median_qi() %>%
    kable()
```

```{r, cache=TRUE}
fit_tt_bayes_transp %>%
    gather_draws(b_Intercept, b_distance_km) %>%
    ggplot(aes(.value)) +
    geom_density() +
    facet_wrap(~ .variable, scales = 'free') +
    labs(x = 'Value',
         title = 'Bayesian Tuesday & Thursday Model w Confounder')
```


# Discussion of the results

# Discussion of the survey/study design

